Detailed Responses to the Referees for \\ ``Beyond Horizontal and Vertical'' \\MS 2023138

### 2024-10-07 doc/JIE/JINDECMS2023138-1Response.tex
Gloria's 10/3 edits

We have created new Section 3.2 that explains our calibration.  We tie our explanation directly to the data inputs, which appear in Table 1, and to the set of equations we use for identification, which are the firm first order conditions in Section 2.  We have considered in Section 3.3 the sensitivity of our results to the share of the outside good/market elasticity, to adding a nesting parameter, and to the size of the bargaining parameter, which we discuss further below.

\item \emph{Be sure to compare simulations that allow for the complex nature of Republic/Santek versus the more vanilla simulations considered in ST (2021). This is important to capture this papers contribution.}

@@

We have added three new sensitivity analyses, at the end of Section 3.3, in order to address this point.  We show how our simulated consumer welfare results are impacted by (1) the extent of substitution outside the market, (2) the use of the logit versus the nested logit, and (3) the bargaining power of retailers vesus wholesalers.  In each of these exercises, we vary the relevant parameter in smooth increments and show how welfare changes relative to the baseline.  The other aspects of the simulations are held fixed, allowing for cleaner comparisons.  We find that our results change in an intuitive manner that we believe will help practitioners understand the impact of the key parameters on the model's conclusions.

@@

    \emph{Note: Footnote 16 states that the antitrust R package contains the code to do the simulations. I could not easily find the relevant parts of the package. It would be great if this machinery was also added to antitrust , but I would like to see a stand-alone replication package.}

     Much of the coding machinery we used has already been incorporated into the next version of the \texttt{antitrust} R package.  A new function, \texttt{vertical.barg}, has been added which uses observed shares, prices and some margins to calibrate demand parameters and simulate the effect of complex mergers. We understand that having a replication package would also be helpful for readers, and we are planning to post a package once the paper has been finalized. Aside from containing all the code and data needed to run all the simulations presented in the paper, the replication package will contain code that allows calibration when downstream costs but not prices are observed.

@@

    \item  \emph{It might make sense if a retailer and manufacturer had meaning in the model. One suggestion is to have $\delta$ be comprised of two components, and manufacturer effect, and a retailer effect, where the retailer effect is perfectly correlated across manufacturer products within a retailer, and the manufacturer effect is perfectly correlated across retailer for the same product. So something like  $\delta_{rw} = \delta_{r} + \delta_{w}$}

    We tried something in this spirit in our nested logit sensitivity exercise for the Republic/Santek simulations.  We did one version where we grouped together all the products that use the same firm for collection and disposal (``integrated'' in Figure 4) and another version where we grouped together all products that use the same landfill (``landfill'' in Figure 4).

@@

    This is a good point.  We have added a footnote to the discussion of that equation (now equation ([8 XXXX])) mentioning the Salinger (1991) result and the Luco and Marshall (2020) paper.

@@

Thank you for this helpful feedback.  We have removed both sets of material, pivoted the draft to focusing on the Republic/Santek merger, which has greatly streamlined the paper.

@@

    We agree that there has been past research on complex mergers, but we also believe that these mergers have been understudied relative to pure vertical and pure horizontal transactions.  Furthermore, we also believe that research showing how a more complicated structural model can be simplified for use by practitioners, keeping in mind the data and time constraints they face, is important.  We have personal experience with being told Crawford et al. (2018) was too complicated to be used as part of the court testimony we were preparing in a merger litigation.

    We have added to the literature review in Section 1 to make clear that there are past papers that have studied complex mergers, including the two you mention above.  We also have added a discussion of earlier theory literature on vertical integration to provide more context for our paper.

@@

@@

    We understand the confusion around this point.  Our meaning of the term agrees with Crawford et al. (2018).  We use the term ``RRC'' in the way that they use it, to mean a shift in bargaining leverage that allows a vertically integrated firm to raise input prices to a downstream rival.  Crawford et al. (2018) refer to this as ``raising rivals' cost.'' See, for example, their discussion in Section 6.2.2, pages 934-936, of their paper.  Footnote 62 in their paper explains that this RRC effect is not motivated by the likelihood that downstream firms will raise their retail prices in response to higher wholesale prices.  The behavior is instead motivated by the possibility that consumers will switch who they purchase from, at fixed prices.  Additional discussion is provided in their footnote 36.  We and Crawford et al. (2018) do not use RRC to mean the incentive modeled by Salop and Scheffman (1983), where a dominant firm chooses an action in order to directly impact downstream retail prices.  We have added footnote [10 XXXX] in the discussion around equation ([9 XXXX]) to clarify this distinction.

@@

    We have removed those simulations from the paper.  All the simulations and calibration we have done rely on the first order conditions of the model in Section 2, to ensure that the resulting prices and market share are consistent with the theory.  We have added a discussion clarifying this point in Section 3.2.

    \item \emph{Finally, I found the calibration exercise in Section 4 hard to follow. At times it reads as if the paper is collecting data. But it is closer to a calibration/simulation exercise motivated by the merger. But there is no estimation and the data is minimal. The theoretical model does not fit the setting in the case. First, the bargaining model is not present nor discussed in the DOJ Lawsuit. Second, the RRC discussed in the DOJ lawsuit refers to foreclosure/exclusionary conduct concerns. This feature could typically has different welfare implications. See, e.g., Ordover et al. (1990) and the references therein. I found the analysis in this section not credible.}

@@

    The DOJ complaint in Republic/Santek at paragraph 47 explains that the DOJ was alleging harm due to increased upstream disposal prices to rival downstream collection firms, not due to total exclusion or foreclosure of these collection firms.  Specifically, the complaint states that the merged firm would have an ``incentive to raise the MSW disposal costs of independent haulers because Republic---no longer confronting competition from Santek in SCCW collection---would capture more of the business lost by independent haulers in the SCCW collection market.''  We agree that the complaint does not put this story in terms of bargaining, but the incentive discussed in this passage is very similar to the mechanism that leads to increased bargaining leverage and hence higher wholesale prices in our model.  The knowledge that downstream customers have fewer independent rivals from which to purchase collection is what allows the merged firm to negotiate for a higher upstream disposal rate.

>>>>

We have created a new Section 3.2 that explains our calibration.  We tie our explanation directly to the data inputs, which appear in Table 1, and to the set of equations we use for identification, which are the firm first order conditions in Section 2.  We have considered in Section 3.3 the sensitivity of our results to the share of the outside good/market elasticity, to adding a nesting parameter, and to the size of the bargaining parameter, which we discuss further below.

\item \emph{Be sure to compare simulations that allow for the complex nature of Republic/Santek versus the more vanilla simulations considered in ST (2021). This is important to capture this paper's contribution.}

@@

We have added three new sensitivity analyses, at the end of Section 3.3, in order to address this point.  We show how our simulated consumer welfare results are impacted by (1) the extent of substitution outside the market, (2) the use of the logit versus the nested logit, and (3) the bargaining power of retailers versus wholesalers.  In each of these exercises, we vary the relevant parameter in smooth increments and show how welfare changes relative to the baseline.  The other aspects of the simulations are held fixed, allowing for cleaner comparisons.  We find that our results change in an intuitive manner that we believe will help practitioners understand the impact of the key parameters on the model's conclusions.

@@

    \emph{Note: Footnote 16 states that the antitrust R package contains the code to do the simulations. I could not easily find the relevant parts of the package. It would be great if this machinery was also added to antitrust, but I would like to see a stand-alone replication package.}

     Much of the coding machinery we used has already been incorporated into the next version of the \texttt{antitrust} R package.  A new function, \texttt{vertical.barg}, has been added that uses observed shares, prices, and some margins to calibrate demand parameters and simulate the effect of complex mergers. We understand that having a replication package would also be helpful for readers, and we are planning to post a package once the paper has been finalized. Aside from containing all the code and data needed to run all the simulations presented in the paper, the replication package will contain code that allows calibration when downstream costs rather than prices are observed.

@@

    \item  \emph{It might make sense if a retailer and manufacturer had meaning in the model. One suggestion is to have $\delta$ be comprised of two components, a manufacturer effect, and a retailer effect, where the retailer effect is perfectly correlated across manufacturer products within a retailer, and the manufacturer effect is perfectly correlated across retailers for the same product. So something like  $\delta_{rw} = \delta_{r} + \delta_{w}$.}

    We tried something in this spirit in our nested logit sensitivity exercise for the Republic/Santek simulations.  We did one version where we grouped together all the products that use the same firm for collection and disposal (``integrated'' in Figure 4) and another version where we grouped together all products that use the same landfill (``landfill'' in Figure 4).  The discussion of these results is in Section 3.3.

@@

    This is a good point.  We have added a footnote to the discussion of that equation (now equation (8)) mentioning the Salinger (1991) result and the Luco and Marshall (2020) paper.

@@

Thank you for this helpful feedback.  We have removed both sets of material and pivoted the draft to focusing on the Republic/Santek merger, which has greatly streamlined the paper.

@@

    We agree that there has been past research on complex mergers, but we also believe that these mergers have been understudied relative to pure vertical and pure horizontal transactions.  Furthermore, we also believe that research showing how a more complicated structural model can be simplified for use by practitioners, keeping in mind the data and time constraints they face, is important.  We have personal experience with being told Crawford et al. (2018) was too complicated to be used as part of the court testimony we were helping to prepare in a merger litigation.

    We have added to the literature review in the Introduction to make clear that there are past papers that have studied complex mergers, including the two you mention above.  We also have added a discussion of earlier theory literature on vertical integration to provide more context for our paper.

@@

    We understand that simultaneous pricing and linear fees may not be appropriate for all applications.  We believe they are useful for some industries, like cable programming, where prices are fixed for periods of time and linear fee contracts are common.  We have added a discussion of these limitations in the Introduction.

@@

    We understand the confusion around this point.  Our meaning of the term agrees with Crawford et al. (2018).  We use the term ``RRC'' in the way that they use it, to mean a shift in bargaining leverage that causes a vertically integrated firm to raise input prices to a downstream rival.  Crawford et al. (2018) refer to this as ``raising rivals' cost.'' See, for example, their discussion in Section 6.2.2, pages 934-936, of their paper.  Footnote 62 in their paper explains that this RRC effect is not motivated by the likelihood that downstream firms will raise their retail prices in response to higher wholesale prices.  The behavior is instead motivated by the possibility that consumers will switch who they purchase from, at fixed prices.  Additional discussion is provided in their footnote 36.  We and Crawford et al. (2018) do not use RRC to mean the incentive modeled by Salop and Scheffman (1983), where a dominant firm chooses an action in order to directly impact downstream retail prices.  We have added footnote 10 in the discussion around equation (9) to clarify this distinction.

@@

    We have removed those simulations from the paper.  All the simulations and calibration we have done rely on the first order conditions of the model in Section 2, to ensure that the resulting prices and market shares are consistent with the theory.  We have added a discussion clarifying this point in Section 3.2.

    \item \emph{Finally, I found the calibration exercise in Section 4 hard to follow. At times it reads as if the paper is collecting data. But it is closer to a calibration/simulation exercise motivated by the merger. But there is no estimation and the data is minimal. The theoretical model does not fit the setting in the case. First, the bargaining model is not present nor discussed in the DOJ lawsuit. Second, the RRC discussed in the DOJ lawsuit refers to foreclosure/exclusionary conduct concerns. This feature could typically have different welfare implications. See, e.g., Ordover et al. (1990) and the references therein. I found the analysis in this section not credible.}

@@

    The DOJ complaint in Republic/Santek at paragraph 46 explains that the DOJ was alleging harm due to increased upstream disposal prices to rival downstream collection firms, not just due to the possibility of total exclusion or foreclosure of these collection firms.  Specifically, the complaint states that the merged firm would have an ``incentive to raise the MSW disposal costs of independent haulers because Republic---no longer confronting competition from Santek in SCCW collection---would capture more of the business lost by independent haulers in the SCCW collection market.''  We agree that the complaint does not put this story in terms of bargaining, but the incentive discussed in this passage is very similar to the mechanism that leads to increased bargaining leverage and hence higher wholesale prices in our model.  The knowledge that downstream customers have fewer independent rivals from which to purchase collection is what allows the merged firm to negotiate for a higher upstream disposal rate.
